Data Prep

### 2024-03-29 scripts/00.0_dataPrep.qmd
fixed downsampling function and reduced 8 million rows to 5 million

```{r load-packages}
library(vultureUtils)
library(sf)
library(tidyverse)
library(move)
library(feather)
library(readxl)
library(elevatr)
library(here)
library(furrr)
future::plan(future::multisession, workers = 10) # this will last for the whole R session.
```

@@

@@

```{r load-hospital}
joined <- read_feather(here::here("data", "dataPrep", "joined.feather"))
periods_to_remove <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "periods_to_remove")
```

```{r remove-periods}
removed_periods <- vultureUtils::removeInvalidPeriods(dataset = joined, periodsToRemove = periods_to_remove)
save(removed_periods, file = here("data/dataPrep/removed_periods.Rda"))
```

@@

# Function to downsample the data
subsample <- function(df, idCol = "Nili_id", 
                      timestampCol = "timestamp", mins = 10){
  sub <- df %>%
    mutate(d = lubridate::date(.data[[timestampCol]])) %>% 
    group_by(.data[[idCol]]) %>%
    arrange(timestampCol, .by_group = T) %>%
    mutate(tc = cut(.data[[timestampCol]], 
                    breaks = paste(as.character(mins), 
                                   "min", sep = " "))) %>%
    ungroup() %>%
    group_by(.data[[idCol]], d, tc) %>%
    slice(1) %>%
    ungroup() %>%
    dplyr::select(-c("d", "tc"))
  return(sub)
}

downsampled_10min_forSocial <- map(removed_lfr, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)

>>>>

```{r load-packages}
library(vultureUtils)
library(sf)
library(tidyverse)
library(move)
library(feather)
library(readxl)
library(elevatr)
library(here)
library(furrr)
future::plan(future::multisession, workers = 5) # this will last for the whole R session.

# function to downsample data to an arbitrary number of minutes, assuming df contains only one individual and timestamp column is called "timestamp"
down <- function(df, mins){
  df <- dplyr::arrange(df, timestamp)
  howmuch <- paste(mins, "minute", sep = " ")
  rounded <- lubridate::round_date(df$timestamp, howmuch)
  keep <- !duplicated(rounded)
  return(df[keep,])
}
```

@@

## Downsample to 5-minute frequency
This data is just WAY too huge to handle now that we have the higher-frequency fixes from the summer/fall of 2023. Going to downsample it a bit.

```{r}
# split `joined` by `Nili_id`
lst <- joined %>% group_split(Nili_id)

# apply the function to each of the ~200 individuals
interval_5min <- map(lst, ~down(df = .x, mins = 5), .progress = T)
interval_5min <- purrr::list_rbind(interval_5min)
write_feather(interval_5min, here("data/dataPrep/interval_5min.feather"))
nrow(joined)
nrow(interval_5min) # okay, that removed a lot!
```

@@

```{r load-hospital}
interval_5min <- read_feather(here::here("data/dataPrep/interval_5min.feather"))
periods_to_remove <- read_excel(here::here("data/raw/whoswho_vultures_20230920_new.xlsx"), sheet = "periods_to_remove")
```

```{r remove-periods}
removed_periods <- vultureUtils::removeInvalidPeriods(dataset = interval_5min, periodsToRemove = periods_to_remove)
write_feather(removed_periods, here("data/dataPrep/removed_periods.feather"))
```

@@

```{r}
#| eval: false
downsampled_10min_forSocial <- map(removed_lfr, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)
